---
title: "Testing the Efficacy of the AF647 in Labelling"
date: "2022-08-26"
author: Brady Johnston
fig-align: center
code-fold: false
code-link: true
toc: true
---

# Introduction
In the previous attempt at in labelling in [previous attempt](dye-labelling.qmd) I was unsuccessful in labelling the protein with _any_ amount of AF647. I had tested both 1, 1.5 and 2 times the ratio of AF647 to Cy3, to account for inefficiencies in labelling. In this experiment I conducted labelling with equimolar concentrations of Cy3 and AF647, but also labelled the the protein with either by itself as well. As a control I also conducted the labelling experiment on BSA, which has a free cysteine that is solvent exposed and able to react with maleimide dyes.

```{r}
#| message: false
#| warning: false
library(here)
library(tidyverse)
library(chromr)
```

# Read in the Chromatograms

```{r}
fl <- list.files(here("data/sec"), pattern = "20220826", full.names = TRUE)

dat <- tibble(
  file = fl
)

dat <- dat |> 
  separate(
    col = file, 
    into = c(NA, NA, "sample", "dye1", "dye2"), 
    sep = "_", 
    remove = FALSE
    ) |> 
  mutate(
    dye2 = str_remove(dye2, ".TXT"),
    data = map(file, chromr::chrom_read_quadtech)
    ) |> 
  unnest(data)
dat |> 
  head()
```

# Analytical SEC

Testing the labelling effectiveness, using analytical SEC after incubation for labelling. I had 6 different labelling reactions, with two different proteins. Need to do a bit of cleanup, trimming the volume and assigning the samples.

```{r}
#| fig-height: 8
dat <-
  dat |> 
  group_by(file) |> 
  filter(volume > 0.5) |> 
  mutate(
    value = value / max(value)
  ) |> 
  filter(!is.na(wl)) |> 
  mutate(
    value = value * 1000,
    dye = case_when(
      dye1 == "Cy3" & dye2 == "AF647" ~ 4, 
      dye1 == "NA" & dye2 == "AF647" ~ 3, 
      dye1 == "Cy3" & dye2 == "NA" ~ 2, 
      TRUE ~ 1
    ), 
    dye = factor(dye, levels = 1:4, labels = c("Protein", "+Cy3" , "+AF647", "+Cy3 +AF647"))
  )

dat |> 
  head()
```


Plotting the adjusted absorbance levels.

```{r}
#| fig-height: 8
#| label: fig-absorbances
#| code-fold: true
#| fig-cap: "Absorbances for the different labelling reactions. Each wavelength measured is a different row, and each sample is a column. The different samples (protein only, Protein + Cy3, Protein + AF647, Protein + Both Dyes) are coloured separtely according to the legend."
dat |> 
ggplot(aes(volume, value, colour = dye)) + 
  geom_line(size = 1, alpha = 0.7) + 
  coord_cartesian(xlim = c(0.5, 4)) +
  scale_colour_manual(
    values = c(
      "gray20", 
      "tomato", 
      "skyblue", 
      "purple"
    )
  ) +
  facet_grid(
    cols = vars(sample), 
    rows = vars(wl)
  ) + 
  theme_bw(base_size = 14) + 
  theme(
    panel.grid = element_line(colour = gray(0.8, 0.1)), 
    legend.position = "top"
  ) + 
  labs(
    colour = "Labelling Reaction", 
    x = "Volume (mL)", 
    y = "Absorbance (mAU)"
  ) -> plt
plt
```

It seems that while on BSA with both dyes present, there is still some labelling effectiveness, but that this is not the case with the PPR protein.

Either BSA or dsnPPR10-C2, which either Cy3 only, AF647 only, or both Cy3 & AF647. It seems that with BSA and both dyes present, they labelled similarly, but in the case of dsnPPR10, when there was Cy3 added it basically didn't label at all.

The presence of Cy3 is outcompeting the labelling of the dsnPPR10-C2, but not BSA?


# Normalised Concentrations

To properly compare the presence of either dye, the 'relative concentration' was calculated according to the Beer-Lambert law (@eq-beer-lambert ).

$$
A = \epsilon b C
$$ {#eq-beer-lambert}

Where $A$ is the absorbance ($AU$), $\epsilon$ is the extinction coefficient ($cm^{-1}M^{-1}$), $b$ is the path length ($cm$) and $C$ is the concentration ($molL^{-1}$). The path length is the same for both and so can be set to 1 for calculating the relative concentrations with @eq-relative-concentration.

$$
C = \frac{A}{\epsilon}
$$ {#eq-relative-concentration}



```{r}
#| label: fig-normalised-conc
#| fig-height: 8
#| code-fold: true
#| fig-cap: Normalised concentration of measured dyes. Concentrations were normalised by dividing measured absorbance readings by extinction coefficients and normalising to the highest value for each labelling reaction. 
dat |> 
  ungroup(file) |> 
  # select(-file) |> 
  filter(wl %in% c(550, 650)) |> 
  mutate(
    value = value / if_else(wl == 650, 265000, 150000)
  ) |> 
  # group_by(sample) |> 
  mutate(
    value = value / max(value)
  ) |> 
  ggplot(aes(volume, value, colour = factor(wl))) +
  geom_line(size = 1, alpha = 0.8) + 
  facet_grid(
    cols = vars(sample), 
    rows = vars(dye)
  ) + 
  theme_bw(base_size = 14) + 
  coord_cartesian(ylim = c(0, NA), xlim = c(0.5, 4)) + 
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  scale_colour_discrete(labels = c("Cy3", "AF647")) +
  labs(
    x = "Elution Volume (mL)", 
    y = "Normalised Concentration", 
    colour = "Dye Measured"
  ) + 
  theme(
    legend.position = "top"
  )
```


Plotting relative concentrations shows that the ratios of AF647 to Cy3 is even worse than what is shown just from the absorbance readings.It seems that even just the presence of Cy3 mostly completely out-competes the AF647 for efficacy in labelling of either protein.

{{< include _page_built_on.qmd >}}
